NCAM Explorer

### 2020-03-14 inst/rmd/vis_fit.Rmd
Revamped NCAM explorer

---
title: "NCAM Explorer"
output: flexdashboard::flex_dashboard
always_allow_html: yes
---

<style>
/* list */
li {
line-height: 30px;
}
</style>

@@

cols <- viridis::viridis(4)[-1]
cols_rgb <-  paste0("rgb(", apply(col2rgb(cols), 2, paste, collapse = ","), ")")
fit <- tmp$fit # temp object from vis_fit function
if (!is.null(tmp$comp)) { comp <- tmp$comp }
if (!is.null(tmp$last)) { last <- tmp$last }
est <- as.list(fit$sd.rep, "Est")
sd <- as.list(fit$sd.rep, "Std")
tabs <- tidy_model(fit)
if (!is.null(tmp$comp)) { comp_tabs <- tidy_model(model_list = comp) }
if (!is.null(tmp$last)) { last_tabs <- tidy_model(tmp$last) }
if (!is.null(fit$retro)) { retro_tabs <- tidy_retro(fit) } 
min_year <- min(fit$tmb.data$years)
max_year <- max(fit$tmb.data$years)
ages <- fit$tmb.data$ages
years <- fit$tmb.data$years
years_plus1 <- fit$tmb.data$years_plus1

## helper functions for nice Rmarkdown text
text_range <- function(x) paste(range(x), collapse = "-")
text_c <- function(x) {
  paste0(paste(x[-length(x)], collapse = ", "), ", and ", x[length(x)])
}
text_units <- function(x) {
  if (x >= 1000000) return(paste0(round(x / 1000000), "M"))
  if (x >= 1000) return(paste0(round(x / 1000), "k"))
  if (x < 1000) return(as.character(round(x)))
}

@@

Background
================================================================================

Row {.tabset}
--------------------------------------------------------------------------------

### NCAM

<h1>NCAM: A state-space stock assessment model for Northern cod</h1>

```{r}

ax <- list(
  title = "",
  zeroline = FALSE,
  showline = FALSE,
  showticklabels = FALSE,
  showgrid = FALSE
)

xy <- approx(x = years_plus1, y = fit$rep$abundance + 1000000, n = 50)
plot_ly(x = xy$x, y = xy$y, hoverinfo = "none",
        fillcolor = 'rgba(49, 104, 142, 0.3)',
        type = 'scatter', mode = 'none', fill = 'tozeroy', showlegend = FALSE) %>%
  add_trace() %>%
  layout(xaxis = ax, yaxis = c(list(range = c(0, max(xy$y))), ax),
         margin = list(l = 0, r = 0, b = 0, t = 0, pad = 0)) %>% 
  plotly::config(displayModeBar = FALSE)

@@

### General

- An *integrated* *state-space* population dynamics model
- *Integrated*
    - Uses as much information as possible
- *State-space*
    - Includes process and observation equations + error
    - Population processes are informed by observations from multiple sources
    - Process errors are necessary for realistic stochastic projections

### Data inputs

- Reported landings (`r text_range(years)`)
- Catch-at-age information (`r text_range(years)`)
- Reported landings by month (`r text_range(years)`)
- Offshore research vessel (RV) survey indices (`r text_range(c(1983, max(years)))`)
- Inshore sentinel (SN) gillnet survey indices (`r text_range(fit$tmb.data$SN_years)`)
- Inshore acoustic biomass surveys and age composition data from Smith Sound (SS; `r text_range(fit$tmb.data$SS_years)`)
- Tagging data (`r text_range(years)`)

### Biological inputs

- RV survey data were used fit cohort-based models of:
   - Proportion mature at age
   - Beginning of year and mid-year stock weights at age
- Baseline natural mortality (M) values

```{r}

plot_ly(x = years, y = ages, z = fit$rep$M_matrix, height = 400, width = 500) %>%
  add_surface() %>%
  colorbar(title = "Mshift") %>%
  layout(scene = list(
    xaxis = list(title = "Year"),
    yaxis = list(title = "Age"),
    zaxis = list(title = "Mshift")
  ))

@@

### Partial catches

- User sets lower and upper bounds on total catch weight
- Bounds are used via a censored likelihood

```{r}

plot_ly(x = years, y = fit$tmb.data$LRL, height = 300, width = 500) %>%
  add_lines(name = "Lower", color = I(cols[1])) %>%
  add_lines(y = fit$tmb.data$LRU, name = "Upper", color = I(cols[2])) %>%
  add_lines(y = rep(1, length(years)), color = I("black"),
            line = list(width = 1, dash = "dot"), showlegend = FALSE) %>%
  layout(yaxis = list(title = "Catch multiplier"),
         xaxis = list(title = "Year"))

@@

### Change in Q

- A shift in Q was assumed for the offshore RV survey
- A variable portion of the stock is assumed to occupy Smith Sound between `r text_range(fit$tmb.data$SS_years)`
- Q for the inshore sentinel survey was assumed to follow a random walk
    - Done to account for potential changes in migration patterns

### Tagging data

- Age of tagged fish estimated from length-at-release using a spatiotemporal Von Bertalanffy growth model
- Applies adjustments for tagging mortality, tag loss and reporting rates
- Reporting rates estimated by an external model using the high-reward tag data collected since 1997
- The Baranov catch equation was applied to age-based tag survivors
- Assumes that initial fishing mortality experienced by fish in each experiment are random deviations from the fishing mortality experienced by the whole stock

### References

Cadigan, N. 2015. A state-space stock assessment model for Northern cod, including under-reported catches and variable natural mortality rates. Canadian Journal of Fisheries and Aquatic Sciences 73 (2). NRC Research Press: 296-308.

Cadigan, N. 2016. Updates to a Northern cod (*Gadus morhua*) state-space integrated assessment model. DFO Can. Sci. Advis. Sec. Res. Doc. 2016/022. v + 58 p.

Catch
================================================================================

Row
--------------------------------------------------------------------------------

### Landings

```{r}

plot_landings(fit)

@@

> Y-axes in log scale

Row {.tabset}
--------------------------------------------------------------------------------

### Observed and predicted catch at age

```{r}

d <- tidy_mats(list(obs = fit$tmb.data$catch_prop,
                    pred = fit$rep$pred_prop_catch),
               ages = ages, years = years)
facet_obs_pred(d, by = "age",
               which = c(2, 8, 3, 9, 4, 10, 5, 11, 6, 12, 7, 13), # done to order by column
               nrows = 6, ylab = "Proportion",
               obs_col = cols[2], pred_col = cols[1]) %>%
  layout(yaxis = list(rangemode = "tozero"))

@@

### Residuals by year, cohort, age and expected values

```{r}

d <- tidy_mats(list(res = fit$rep$CA_resid,
                    fit = fit$rep$pred_crl),
               ages = ages[-length(ages)], years = years)
d$cohort <- d$year - d$age
d$rec <- seq(nrow(d))
plot_resid(d, col = cols[1])

@@

> based on continuation ratio logits

### Catch at age surface (proportion)

```{r}
plot_surface(mat = fit$rep$pred_prop_catch, ages = ages, years = years, name = "P")
```

<!-- ### Catch at age surface (numbers) -->

<!-- ```{r} -->
<!-- plot_surface(mat = t(inputs$catch.num), ages = ages, years = years, name = "N") -->
<!-- ``` -->

<!-- > Note: model fits to composition - this plot is included for illustration -->

RV Survey
================================================================================

Row
--------------------------------------------------------------------------------

### RV survey index

```{r}

obs <- colSums(fit$tmb.data$est_wt * fit$tmb.data$RV_index * fit$tmb.data$trawl_units)
pred <- colSums(fit$tmb.data$est_wt * exp(fit$rep$Elog_RV_index) * fit$tmb.data$trawl_units)
obs[obs == 0] <- NA
pred[pred == 0] <- NA
d <- data.frame(year = years, obs = obs, pred = pred)

# ## RV biomass
# obs <- colSums(fit$tmb.data$RV_index * fit$tmb.data$trawl_units * fit$tmb.data$weight)
# pred <- colSums(exp(fit$rep$Elog_RV_index) * fit$tmb.data$trawl_units * fit$tmb.data$weight)
# d <- data.frame(year = years, obs = obs, pred = pred)
# write.table(d, file = "clipboard", sep = "\t", row.names = FALSE, quote = FALSE)

plot_obs_pred(d, obs_col = cols[2], pred_col = cols[1], ylab = "RV survey index",
              showlegend = TRUE, type = "log")

@@

> Y-axes in log scale

### Standardized log residuals

```{r}

d <- tidy_mats(list(res = fit$rep$RV_resid,
                    zero = fit$tmb.data$ind_zero,
                    wt = fit$tmb.data$est_wt),
               ages = ages, years = years)
d <- d[d$wt == 1, ]
bubble_plot(d)

Row {.tabset}
--------------------------------------------------------------------------------

### RV survey index at age

```{r}

d <- tidy_mats(list(obs = fit$tmb.data$RV_index,
                    pred = exp(fit$rep$Elog_RV_index),
                    wt = fit$tmb.data$est_wt),
               ages = ages, years = years)
d$obs[d$wt == 0] <- NA
d$pred[d$wt == 0] <- NA
facet_obs_pred(d, by = "age",
               which = c(2, 8, 3, 9, 4, 10, 5, 11, 6, 12, 7, 13), # done to order by column
               nrows = 6, ylab = "Mean numbers per tow",
               obs_col = cols[2], pred_col = cols[1], type = "log")

> Y-axes in log scale

### Residuals by year, cohort, age and expected values

```{r}

d <- tidy_mats(list(res = fit$rep$RV_resid,
                    keep = fit$tmb.data$ind_zero == 0 & fit$tmb.data$est_wt == 1,
                    fit = fit$rep$Elog_RV_index),
               ages = ages, years = years)
d <- d[d$keep, ]
d$cohort <- d$year - d$age
d$rec <- seq(nrow(d))
plot_resid(d, col = cols[1])

SN Survey
================================================================================

Row
--------------------------------------------------------------------------------

### Sentinel gillnet survey index

```{r}

obs <- colSums(fit$tmb.data$SN_GN_index)
pred <- colSums(exp(fit$rep$Elog_SN_GN_index))
d <- data.frame(year = fit$tmb.data$SN_years, obs = obs, pred = pred)

plot_obs_pred(d, obs_col = cols[2], pred_col = cols[1], ylab = "Sentinel survey index",
              showlegend = TRUE, type = "log")

> Y-axes in log scale

### Standardized log residuals

```{r}

d <- tidy_mat(fit$rep$log_SN_GN_resid,
              ages = fit$tmb.data$SN_ages,
              years = fit$tmb.data$SN_years, value = "res")
bubble_plot(d)

Row {.tabset}
--------------------------------------------------------------------------------

### Sentinel gillnet survey index at age

```{r}

d <- tidy_mats(list(obs = fit$tmb.data$SN_GN_index,
                    pred = exp(fit$rep$Elog_SN_GN_index)),
               ages = fit$tmb.data$SN_ages, years = fit$tmb.data$SN_years)
facet_obs_pred(d, by = "age",
               which = c(3, 7, 4, 8, 5, 9, 6, 10), # done to order by column
               nrows = 4, ylab = "Sentinel survey index",
               obs_col = cols[2], pred_col = cols[1], type = "log")

> Y-axes in log scale

### Residuals by year, cohort, age and expected values

```{r}

d <- tidy_mats(list(res = fit$rep$log_SN_GN_resid,
                    fit = fit$rep$Elog_SN_GN_index),
               ages = fit$tmb.data$SN_ages,
               years = fit$tmb.data$SN_years)
d$cohort <- d$year - d$age
d$rec <- seq(nrow(d))
plot_resid(d, col = cols[1])

SS Survey
================================================================================

Row
--------------------------------------------------------------------------------

### Smith sound acoustic biomass estimates

```{r}

d <- data.frame(year = fit$tmb.data$SS_bms_years,
                obs = fit$tmb.data$SSbms,
                pred = fit$rep$ESSbms)
plot_obs_pred(d, obs_col = cols[2], pred_col = cols[1], ylab = "Biomass (t)",
              showlegend = TRUE)

### Standardized log residuals

```{r}

d <- tidy_mat(fit$rep$SS_agecomps_resid_std,
              ages = fit$tmb.data$SS_ages,
              years = fit$tmb.data$SS_age_years, value = "res")
bubble_plot(d)

Row {.tabset}
--------------------------------------------------------------------------------

### Smith Sound trawl numbers at age

```{r}

d <- tidy_mats(list(obs = fit$tmb.data$SS_agecomps,
                    pred = fit$rep$ESS_agecomps),
               ages = fit$tmb.data$SS_ages,
               years = fit$tmb.data$SS_age_years)
facet_obs_pred(d, by = "age",
               which = c(2, 8, 3, 9, 4, 10, 5, 11, 6, 12, 7, 13), # done to order by column
               nrows = 6, ylab = "Numbers caught", rangemode = "tozero",
               obs_col = cols[2], pred_col = cols[1])

### Residuals by year, cohort, age and expected values

```{r}

d <- tidy_mats(list(res = fit$rep$SS_agecomps_resid_std,
                    fit = fit$rep$ESS_agecomps),
               ages = fit$tmb.data$SS_ages,
               years = fit$tmb.data$SS_age_years)
d$cohort <- d$year - d$age
d$rec <- seq(nrow(d))
plot_resid(d, col = cols[1])

Tagging {data-orientation=rows}
================================================================================

```{r}

## Prep tag data and fits
tag_dat <- data.frame(
  exp = c(fit$tmb.data$old_tag_exp, fit$tmb.data$tag_exp),
  n_rel = c(fit$tmb.data$old_N_rel, fit$tmb.data$N_rel),
  rel_year = c(fit$tmb.data$old_rel_year, fit$tmb.data$rel_year) + min(years),
  cap_year = c(fit$tmb.data$old_cap_year, fit$tmb.data$cap_year) + min(years),
  age = c(fit$tmb.data$old_tag_age, fit$tmb.data$tag_age),
  pred = c(fit$rep$old_tag_report, fit$rep$tag_report),
  obs = c(fit$tmb.data$old_N_cap, fit$tmb.data$N_cap),
  res = c(fit$rep$old_tag_std_resid, fit$rep$tag_std_resid),
  period = c(rep("pre 1996", length(fit$tmb.data$old_tag_exp)),
             rep("post 1996", length(fit$tmb.data$tag_exp)))
)
tag_dat$delay <- tag_dat$cap_year - tag_dat$rel_year
tag_dat$text <- paste("exp:", tag_dat$exp, "</br>",
                      "age:", tag_dat$age, "</br>",
                      "delay:", tag_dat$delay, "</br>")

Udev <- data.frame(
  exp = c(unique(fit$tmb.data$old_tag_exp), unique(fit$tmb.data$tag_exp)),
  Udev = c(fit$rep$old_Udev, fit$rep$Udev)
)

Fay <- tidy_mat(fit$rep$F_matrix, ages = ages, years = years, value = "Fay")
names(Fay)[names(Fay) == "year"] <- "rel_year"

tag_dat <- merge(tag_dat, Udev, by = "exp")
tag_dat <- merge(tag_dat, Fay, by = c("rel_year", "age"))
tag_dat$Fsay <- tag_dat$Fay * exp(tag_dat$Udev)

## Share data across multiple plots
d <- crosstalk::SharedData$new(tag_dat, ~exp, "Select an experiment")
base <- plot_ly(d, symbol = ~period, colors = cols[1:2], symbols = c(16, 1)) %>% group_by(exp)

Row {data-height=700 .tabset}
--------------------------------------------------------------------------------

### Observed and predicted tag recaptures by experimemnt

```{r}

y_ticktext <- c(50, 500, pretty(c(0, 6000), 4))
y_tickvals <- sqrt(y_ticktext)

rel_plot <- base %>%
  filter(rel_year == cap_year) %>%
  group_by(exp) %>%
  summarise(rel_year = unique(rel_year), tot_rel = sum(n_rel), period = unique(period)) %>%
  add_markers(x = ~rel_year, y = ~sqrt(tot_rel), color = I(cols[1]),
              text = ~paste("exp:", exp, "</br>", "N:", tot_rel)) %>%
  layout(shapes = list(type = "line", x0 = 1996, x1 = 1996,
                       y0 = 0, y1 = 1000, yref = "y",
                       line = list(width = 1, dash = "dot")),
         yaxis = list(title = "N released",
                      tickvals = y_tickvals,
                      ticktext = y_ticktext,
                      range = range(y_tickvals)),
         xaxis = list(title = "Year"))

y_ticktext <- c(5, 20, pretty(c(0, 500), 10))
y_tickvals <- sqrt(y_ticktext)

cap_plot <- base %>%
  group_by(exp, rel_year, cap_year) %>%
  summarise(sum_pred = sum(pred), sum_obs = sum(obs),
            period = unique(period)) %>%
  group_by(exp) %>%
  add_markers(x = ~cap_year, y = ~sqrt(sum_obs), color = I(cols[1]), name = "",
              showlegend = FALSE, text = ~paste("exp:", exp, "</br>", "observed N:", sum_obs)) %>%
  add_lines(x = ~cap_year, y = ~sqrt(sum_pred), color = I(cols[2]), name = "Predicted",
            alpha = 0.7, showlegend = FALSE, symbol = NULL,
            text = ~paste("exp:", exp, "</br>", "predicted N:", round(sum_pred))) %>%
  layout(shapes = list(type = "line", x0 = 1996, x1 = 1996,
                       y0 = 0, y1 = 1000, yref = "y2",
                       line = list(width = 1, dash = "dot")),
         yaxis = list(title = "N recaptured",
                      tickvals = y_tickvals,
                      ticktext = y_ticktext,
                      range = range(y_tickvals)),
         xaxis = list(title = "Year"))

subplot(rel_plot, cap_plot, shareX = TRUE, titleX = TRUE, titleY = TRUE,
        nrows = 2, heights = c(0.3, 0.7))

> Y-axis is in square root scale

<!-- ### Animation -->

<!-- ```{r} -->

<!-- sum_d <- tag_dat %>% -->
<!--   group_by(exp, cap_year) %>% -->
<!--   summarise(sum_pred = sum(pred), sum_obs = sum(obs)) -->

<!-- sum_d %>%  -->
<!--   group_by(exp) %>%  -->
<!--   plot_ly(x = ~cap_year, frame = ~exp) %>% -->
<!--   add_markers(y = ~sqrt(sum_obs), color = I(cols[1]), name = "Observed", -->
<!--               text = ~paste("exp:", exp, "</br>", "observed N:", sum_obs)) %>% -->
<!--   add_lines(y = ~sqrt(sum_pred), color = I(cols[2]), name = "Predicted", -->
<!--             text = ~paste("exp:", exp, "</br>", "predicted N:", round(sum_pred))) %>% -->
<!--   animation_opts(1000, transition = 500, easing = "exp", redraw = FALSE) %>% -->
<!--   layout(yaxis = list(title = "N recaptured", -->
<!--                       tickvals = y_tickvals, -->
<!--                       ticktext = y_ticktext, -->
<!--                       range = range(y_tickvals)), -->
<!--          xaxis = list(title = "Year")) -->

<!-- ``` -->

Row
--------------------------------------------------------------------------------

### Tag reporting rate

```{r}

cov_mat <- Matrix::as.matrix(fit$tmb.data$SRR_cov_matrix)
rr <- data.frame(year = fit$tmb.data$SRR_years,
                 est = fit$tmb.data$log_SRR_est,
                 sd = sqrt(diag(cov_mat)))
rr$lwr <- rr$est - 1.96 * rr$sd
rr$upr <- rr$est + 1.96 * rr$sd
rr <- trans_est(rr)
rr$model <- "Random Walk"

ncam_rr <- data.frame(year = fit$tmb.data$SRR_years,
                      est = est$log_SRR, sd = sd$log_SRR)
ncam_rr$lwr <- ncam_rr$est - 1.96 * ncam_rr$sd
ncam_rr$upr <- ncam_rr$est + 1.96 * ncam_rr$sd
ncam_rr <- trans_est(ncam_rr)
ncam_rr$model <- "NCAM"

rr <- rbind(rr, ncam_rr)

plot_trend(rr, group = "model", col = cols, ylab = "Reporting rates", annotate = TRUE)

### Initial F deviations for tagging experiments

```{r}

base %>%
  distinct(rel_year, Udev, exp, period) %>%
  add_markers(x = ~rel_year, y = ~Udev, text = ~exp, color = I(cols[1]),
              showlegend = FALSE) %>%
  layout(shapes = list(type = "line", x0 = 1996, x1 = 1996,
                       y0 = min(tag_dat$Udev), y1 = max(tag_dat$Udev),
                       line = list(width = 1, dash = "dot")),
         yaxis = list(title = "Initial F deviation"),
         xaxis = list(title = "Year"))

### Observed vs. predicted recaptures

```{r}

x_ticktext <- c(5, pretty(tag_dat$pred, 10))
y_ticktext <- c(5, pretty(tag_dat$obs, 10))
x_tickvals <- sqrt(x_ticktext)
y_tickvals <- sqrt(y_ticktext)

base %>%
  add_markers(x = ~sqrt(pred), y = ~sqrt(obs),
              text = ~text, color = I(cols[1]), showlegend = FALSE) %>%
  layout(shapes = list(type = "line", x0 = 0, x1 = max(sqrt(c(tag_dat$pred, tag_dat$obs))),
                       y0 = 0, y1 = max(sqrt(c(tag_dat$pred, tag_dat$obs))),
                       line = list(width = 1)),
         xaxis = list(title = "Predicted number of recaptures",
                      tickvals = x_tickvals,
                      ticktext = x_ticktext),
         yaxis = list(title = "Observed number of recaptures",
                      tickvals = y_tickvals,
                      ticktext = y_ticktext))

> Axes are in square root scale

### Standardized residuals vs. predicted recaptures

```{r}

y_ticktext <- c(5, pretty(tag_dat$res, 10))
y_tickvals <- sqrt(x_ticktext)

base %>%
  add_markers( x = ~sqrt(pred), y = ~res,
               text = ~text, color = I(cols[1]), showlegend = FALSE) %>%
  layout(xaxis = list(title = "Predicted number of recaptures",
                      tickvals = x_tickvals,
                      ticktext = x_ticktext),
         yaxis = list(title = "Standardized residuals"))

> X-axis is in square root scale

Catchability
================================================================================

Row {.tabset}
--------------------------------------------------------------------------------

### Fishery selectivity

```{r}

S <- apply(fit$rep$F_matrix, 2, function(x) x / max(x))
plot_surface(S, ages = ages, years = years, name = "S")

Row {.tabset}
--------------------------------------------------------------------------------

### RV survey catchability

```{r}

q <- exp(fit$rep$log_RV_Qa + log(fit$tmb.data$trawl_units))
plot_surface(q, ages = ages, years = years, name = "q")

### Proportion of the stock in Smith Sound

```{r}
p <- matrix(0, nrow = length(ages), ncol = length(years),
            dimnames = list(ages, years))
ss_p <- fit$rep$SSD
ss_years <- fit$tmb.data$SS_years
ss_ages <- fit$tmb.data$SS_ages
rownames(ss_p) <- ss_ages
colnames(ss_p) <- ss_years
p[as.character(ss_ages), as.character(ss_years)] <- ss_p
plot_surface(p, ages = ages, years = years, name = "p")

### Sentinel gillnet survey catchability

```{r}

log_qy <- t(replicate(length(fit$tmb.data$SN_ages), fit$rep$SN_GN_Qy))
log_qa <- fit$rep$log_SN_GN_Qa
q <- exp(log_qy + log_qa)
plot_surface(q, ages = fit$tmb.data$SN_ages,
             years = fit$tmb.data$SN_years,
             name = "q")

Trends
================================================================================

Row {.tabset}
--------------------------------------------------------------------------------

### Stock status

```{r}

## Note: biomass is in tonnes and numbers are in thousands
bms <- rbind(data.frame(metric = "SSB", tabs$log_ssb),
             data.frame(metric = "Ages 2+", tabs$log_biomass))
abn <- rbind(data.frame(metric = "Age 2", tabs$log_recruitment),
             data.frame(metric = "Ages 2+", tabs$log_abundance))

bms_plot <- trans_est(bms) %>%
  plot_trend(ylab = "Biomass (t)", group = "metric", col = cols,
             annotate = TRUE)

abn_plot <- trans_est(abn, scale = 0.001) %>%
  plot_trend(ylab = "Abundance", group = "metric", col = cols,
             annotate = TRUE)

blim_plot <- trans_est(tabs$log_ssb_lrp) %>%
  plot_trend(ylab = "SSB/Blim", annotate = TRUE, col = cols) %>%
  add_segments(x = min_year, xend = max_year + 1, y = 1, yend = 1, color = I("black"),
               linetype = I(3), name = "Blim", text = NULL, hoverinfo = "none") %>%
  add_text(x = max_year + 1.2, y = 1, text = paste(round(fit$rep$blim / 1000), "kt"),
           color = I("black"), textposition = "right", hoverinfo = "none")

subplot(abn_plot, bms_plot, blim_plot, nrows = 3, shareX = TRUE, titleY = TRUE)

### Abundance at age

```{r}
plot_surface(fit$rep$N_matrix * 1000, ages = ages, years = years_plus1, name = "N")
```

### Biomass at age

```{r}
plot_surface(fit$rep$biomass_matrix, ages = ages, years = years_plus1, name = "Biomass")
```

### SSB at age

```{r}
plot_surface(fit$rep$ssb_matrix, ages = ages, years = years_plus1, name = "SSB")
```

Row {.tabset}
--------------------------------------------------------------------------------

### Mortality Rates

```{r}

f <- data.frame(metric = "Ages 5+", tabs$log_aveF)
z <- data.frame(metric = "Ages 5+", tabs$log_aveZ)
m <- data.frame(metric = "Ages 5+", tabs$log_aveM)

z_plot <- trans_est(z) %>%
  plot_trend(ylab = "Average Z", group = "metric",
             annotate = TRUE, col = cols)

m_plot <- trans_est(m) %>%
  plot_trend(ylab = "Average M", group = "metric",
             annotate = TRUE, col = cols)

f_plot <- trans_est(f) %>%
  plot_trend(ylab = "Average F", group = "metric",
             annotate = TRUE, col = cols)

subplot(z_plot, m_plot, f_plot, nrows = 3, shareX = TRUE, titleY = TRUE)

### Z at age

```{r}
plot_surface(fit$rep$Z_matrix, ages = ages, years = years, name = "Z")
```

### M at age

```{r}
plot_surface(fit$rep$Mpe_matrix, ages = ages, years = years, name = "M")
```

### F at age

```{r}
plot_surface(fit$rep$F_matrix, ages = ages, years = years, name = "F")
```

### Z, M and F

```{r}
plot_zmf_surface(z = fit$rep$Z_matrix, m = fit$rep$Mpe_matrix, f = fit$rep$F_matrix,
                 ages = ages, years = years)
```

> Z = purples, M = blues, F = reds

### Process error

```{r}

pe <- fit$rep$pe
plot_surface(pe, years = c(years, fit$tmb.data$proj_years)[-1], ages = 2:8,
             name = "pe", col = RColorBrewer::brewer.pal(9, "RdBu")) %>%
  add_surface(matrix(0, nrow = nrow(pe), ncol = ncol(pe)), showscale = FALSE,
              color = I("black"), opacity = 0.8)

### Process error (bubble plot)

```{r}
pe <- tidy_mat(fit$rep$pe, years = c(years, fit$tmb.data$proj_years)[-1],
               ages = 2:8, value = "res")
bubble_plot(pe)
```

Retro {`r if (is.null(fit$retro)) ".hidden"`}
================================================================================

Row
--------------------------------------------------------------------------------

### Stock status

```{r, eval = !is.null(fit$retro)}

bms_plot <- trans_est(retro_tabs$log_ssb) %>%
  plot_retro(ylab = "SSB (t)", lwd = 1, col = cols[1])

abn_plot <- trans_est(retro_tabs$log_recruitment, scale = 0.001) %>%
  plot_retro(ylab = "Recruitment (age 2)", lwd = 1, col = cols[1])

blim_plot <- trans_est(retro_tabs$log_ssb_lrp) %>%
  plot_retro(ylab = "SSB/Blim", lwd = 1, col = cols[1]) %>%
  add_segments(x = min_year, xend = max_year + 1, y = 1, yend = 1, color = I("black"),
               linetype = I(3), name = "Blim", text = NULL, hoverinfo = "none")

subplot(abn_plot, bms_plot, blim_plot, nrows = 3, shareX = TRUE, titleY = TRUE)

Row
--------------------------------------------------------------------------------

### Mortality Rates

```{r, eval = !is.null(fit$retro)}

z_plot <- trans_est(retro_tabs$log_aveZ) %>%
  plot_retro(ylab = "Average Z (ages 5+)", lwd = 1, col = cols[1])

m_plot <- trans_est(retro_tabs$log_aveM) %>%
  plot_retro(ylab = "Average M (ages 5+)", lwd = 1, col = cols[1])

f_plot <- trans_est(retro_tabs$log_aveF) %>%
  plot_retro(ylab = "Average F (ages 5+)", lwd = 1, col = cols[1])

subplot(z_plot, m_plot, f_plot, nrows = 3, shareX = TRUE, titleY = TRUE)

Comps {`r if (is.null(tmp$comp)) ".hidden"`}
================================================================================

Row
--------------------------------------------------------------------------------

### Stock status

```{r, eval = !is.null(tmp$comp)}

abn_plot <- trans_est(comp_tabs$log_recruitment, scale = 0.001) %>% 
  plot_comps(ylab = "Recruitment (age 2)", col = cols)

bms_plot <- trans_est(comp_tabs$log_ssb) %>% 
  plot_comps(ylab = "SSB (t)", col = cols, showlegend = FALSE)

blim_plot <- trans_est(comp_tabs$log_ssb_lrp) %>%
  plot_comps(ylab = "SSB/Blim", col = cols, showlegend = FALSE) %>%
  add_segments(x = min_year, xend = max_year + 1, y = 1, yend = 1, color = I("black"),
               linetype = I(3), name = "Blim", text = NULL, hoverinfo = "none", showlegend = FALSE)

subplot(abn_plot, bms_plot, blim_plot, nrows = 3, shareX = TRUE, titleY = TRUE)

Row
--------------------------------------------------------------------------------

### Mortality Rates

```{r, eval = !is.null(tmp$comp)}

z_plot <- trans_est(comp_tabs$log_aveZ) %>%
  plot_comps(ylab = "Average Z (ages 5+)", col = cols)

m_plot <- trans_est(comp_tabs$log_aveM) %>%
  plot_comps(ylab = "Average M (ages 5+)", col = cols, showlegend = FALSE)

f_plot <- trans_est(comp_tabs$log_aveF) %>%
  plot_comps(ylab = "Average F (ages 5+)", col = cols, showlegend = FALSE)

subplot(z_plot, m_plot, f_plot, nrows = 3, shareX = TRUE, titleY = TRUE)

Productivity
================================================================================

Row {.tabset}
--------------------------------------------------------------------------------

### Recruitment

```{r}
trans_est(tabs$log_recruitment, scale = 0.001) %>%
  plot_trend(ylab = "Recruitment", col = cols[1])
```

### SSB

```{r}
trans_est(tabs$log_ssb) %>%
  plot_trend(ylab = "SSB (t)", col = cols[1])
```

### RPS

```{r}
trans_est(tabs$log_rps) %>%
  filter(year >= min_year + min(ages)) %>% 
  plot_trend(ylab = "RPS (thousand / tonne)", col = cols[1]) %>% 
  layout(xaxis = list(range = c(min_year, max_year + 1)))
```

### SSB x Recruitment

```{r}

ssb <- trans_est(tabs$log_ssb)
rec <- trans_est(tabs$log_recruitment, scale = 0.001)
rps <- trans_est(tabs$log_rps)
names(ssb) <- paste0("ssb_", names(ssb))
names(rec) <- paste0("rec_", names(rec))
names(rps) <- paste0("rps_", names(rps))
ssb$year <- ssb$ssb_year + min(ages)
rec$year <- rec$rec_year
rps$year <- rps$rps_year
sr <- merge(ssb, rec, by = "year")
sr <- merge(sr, rps, by = "year")
sr$lab <- paste0("Recruitment year: ", sr$rec_year, "<br>SSB year: ", sr$ssb_year)

plot_xy_errorbar(sr, x = ~ssb_est, y = ~rec_est,
                 x_lwr = ~ssb_lwr, x_upr = ~ssb_upr,
                 y_lwr = ~rec_lwr, y_upr = ~rec_upr,
                 text = ~lab, xlab = "SSB (t)", ylab = "Recruitment",
                 col = cols[1])

### SSB x RPS

```{r}
p <- plot_xy_errorbar(sr, x = ~ssb_est, y = ~rps_est,
                 x_lwr = ~ssb_lwr, x_upr = ~ssb_upr,
                 y_lwr = ~rps_lwr, y_upr = ~rps_upr,
                 text = ~lab, xlab = "SSB (t)", ylab = "RPS (thousand / tonne)",
                 col = cols[1])
p
```

### log(SSB) x log(RPS)

```{r}
p %>% layout(yaxis = list(type = "log"),
             xaxis = list(type = "log"))
```

```{r}

>>>>

---
title: "NCAM Explorer"
output: 
  flexdashboard::flex_dashboard:
    css: style.css
---

@@

## https://projects.susielu.com/viz-palette?colors=[%22#243a68%22,%22#6a8b3e%22,%22#ff9800%22,%22#f44336%22,%22#9e9e9e%22]&backgroundColor=%22white%22&fontColor=%22black%22&mode=%22normal%22
## https://material.io/resources/color/#!/?view.left=0&view.right=0&primary.color=243A68&secondary.color=9E9E9E

cols <- c("#243a68", "#6a8b3e", "#ff9800", "#f44336", "#9e9e9e")
cols_rgb <-  paste0("rgb(", apply(col2rgb(cols), 2, paste, collapse = ","), ")")
obs_col <- cols[5]
pred_col <- pos_col <- cols[1]
neg_col <- cols[4]

fit <- tmp$fit # temp object from vis_fit function
if (!is.null(tmp$comp)) { comp <- tmp$comp }
if (!is.null(tmp$last)) { last <- tmp$last }
est <- as.list(fit$sd.rep, "Est")
sd <- as.list(fit$sd.rep, "Std")
est_lwr <- lapply(seq_along(est), function(i) est[[i]] - qnorm(0.975) * sd[[i]])
est_upr <- lapply(seq_along(est), function(i) est[[i]] + qnorm(0.975) * sd[[i]])
names(est_lwr) <- names(est_upr) <- names(est)
tabs <- tidy_model(fit)
if (!is.null(tmp$comp)) { comp_tabs <- tidy_model(model_list = comp) }
if (!is.null(tmp$last)) { last_tabs <- tidy_model(tmp$last) }
if (!is.null(fit$retro)) { retro_tabs <- tidy_retro(fit) } 
min_year <- min(fit$tmb.data$years)
max_year <- max(fit$tmb.data$years)
ages <- fit$tmb.data$ages
years <- fit$tmb.data$years
years_plus1 <- fit$tmb.data$years_plus1

## helper functions for nice Rmarkdown text
text_range <- function(x) paste(range(x), collapse = "-")
text_c <- function(x) {
  paste0(paste(x[-length(x)], collapse = ", "), ", and ", x[length(x)])
}
text_units <- function(x) {
  ifelse(x >= 1000000000, paste0(round(x / 1000000000), "B"),
         ifelse(x >= 1000000, paste0(round(x / 1000000), "M"),
         ifelse(x >= 1000, paste0(round(x / 1000), "k"), 
         as.character(round(x)))))
}

## Function for finding local maxima
## from https://github.com/stas-g/findPeaks
find_peaks <- function (x, m = 3){
  shape <- diff(sign(diff(x, na.pad = FALSE)))
  pks <- sapply(which(shape < 0), FUN = function(i){
    z <- i - m + 1
    z <- ifelse(z > 0, z, 1)
    w <- i + m + 1
    w <- ifelse(w < length(x), w, length(x))
    if(all(x[c(z : i, (i + 2) : w)] <= x[i + 1])) return(i + 1) else return(numeric(0))
  })
  pks <- unlist(pks)
  pks
}

@@

```{r background, child = "background.Rmd"}

@@

```{r fishery, child = "fishery.Rmd"}

@@

```{r surveys, child = "surveys.Rmd"}

@@

```{r tagging, child = "tagging.Rmd"}

@@

```{r trends, child = "trends.Rmd"}

@@

```{r projections, child = "projections.Rmd"}

@@

```{r tables, child = "tables.Rmd"}

@@
